infer: An R package for tidyverse-friendly statistical inference

### 2021-06-25 figs/paper/paper.tex
mine's edits

\texttt{infer} implements an expressive grammar to perform statistical
inference that adheres to the \texttt{tidyverse} design framework
(Wickham et al., 2019). Rather than providing methods for specific
statistical tests, this package consolidates the principles that are
shared among common hypothesis tests into a set of four main verbs
(functions), supplemented with many utilities to visualize and extract
value from their outputs.

@@

\texttt{infer} is an R package for randomization-based hypothesis
testing, naturalizing an intuitive understanding of statistical
inference via a unified and expressive grammar. Four functions provide
functionality encompassing a large swath of basic frequentist
statistical inference, abstracting away details of specific tests and
shifting the focus of the analyst to the observed data and the processes
that generated it. Such a grammar lends itself to applications in
teaching, data pedagogy research, applied scientific research, and
advanced predictive modeling. For one, the principled approach of the
package has made it an especially good fit for teaching introductory
statistics (Baumer, Bray, Çetinkaya-Rundel, \& Hardin, 2020; Ismay \&
Kim, 2019) and research in data pedagogy (Çetinkaya-Rundel \& Ellison,
2021; Fergusson \& Pfannkuch, 2021; Loy, 2021). Further, the package has
already seen usage in a number of published scientific applications (Ask
et al., 2021; Fallon \& Hinds, 2021; McLean et al., 2021). Finally, the
package integrates with the greater tidymodels collection of packages, a
burgeoning software ecosystem for tidyverse-aligned predictive modeling
used across many modern research and industrial applications (Kuhn \&
Wickham, 2020). To date, the package has been downloaded more than
400,000 times.

@@

\begin{itemize}
\tightlist
\item
  \texttt{specify()} allows the analyst to specify the variable, or
  relationship between variables, that they are interested in.
\item
  \texttt{hypothesize()} allows the analyst to declare the null
  hypothesis.
\item
  \texttt{generate()} allows the analyst to generate data reflecting the
  null hypothesis.
\item
  \texttt{calculate()} allows the analyst to calculate summary
  statistics, either from

  \begin{itemize}
  \tightlist
  \item
    the observed data, to form the observed test statistic.
  \item
    data \texttt{generate()}d to reflect the null hypothesis, to form
    the randomization-based null distribution of test statistics.
  \end{itemize}
\end{itemize}

As such, the ultimate output of an infer pipeline using these four
functions is generally an \emph{observed statistic} or \emph{null
distribution} of test statistics. These four functions are thus
supplemented with many utilities to visualize and extract value from
their outputs.

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{shade\_p\_value()} situates the observed statistic in the
    null distribution, shading the region as or more extreme.
  \item
    \texttt{shade\_confidence\_interval()} situates the confidence
    interval region in the null distribution, shading the region within
    the bounds.
  \end{itemize}
\item
  \texttt{get\_p\_value()} calculates a p-value via the juxtaposition of
  the test statistic and null distribution.
\item
  \texttt{get\_confidence\_interval()} calculates a confidence interval
  via the juxtaposition of the test statistic and null distribution.
\end{itemize}

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though do not provide methods for fitting models,
describing null distributions, or calculating summary statistics from
tabular data. \texttt{statsexpressions}, and adjacent packages in the
\texttt{easystats} ecosystem, implement wrappers with consistent
interfaces for theory-based hypothesis tests. At a higher level, the
structure of each of these packages is defined by model types, where
each model type has its own associated function and/or model object. In
contrast, \texttt{infer} is structured around four functions, situating
statistics and model types within a more abstracted grammar.
\texttt{infer}'s initial release on the Comprehensive R Archive Network
predated that of each of the packages mentioned above, with the
exception of \texttt{broom}, a package situated in the same
\texttt{tidymodels} ecosystem (R Core Team, 2021).

>>>>

\texttt{infer} implements an expressive grammar to perform statistical
inference that adheres to the \texttt{tidyverse} design framework
(Wickham et al., 2019). Rather than providing methods for specific
statistical tests, this package consolidates the principles that are
shared among common hypothesis tests and confidence intervals into a set
of four main verbs (functions), supplemented with many utilities to
visualize and extract value from their outputs.

@@

\texttt{infer} is an R package for randomization-based hypothesis
testing, naturalizing an intuitive understanding of statistical
inference via a unified and expressive grammar. Four functions provide
functionality encompassing a large swath of basic frequentist
statistical inference, abstracting away details of specific tests and
shifting the focus of the analyst to the observed data and the processes
that generated it. Such a grammar lends itself to applications in
teaching, data pedagogy research, applied scientific research, and
advanced predictive modeling. For one, the principled approach of the
package has made it an especially good fit for teaching introductory
statistics and data science (Baumer, Bray, Çetinkaya-Rundel, \& Hardin,
2020; Çetinkaya-Rundel \& Ellison, 2021; Ismay \& Kim, 2019) and
research in data pedagogy (Fergusson \& Pfannkuch, 2021; Loy, 2021).
Further, the package has already seen usage in a number of published
scientific applications (Ask et al., 2021; Fallon \& Hinds, 2021; McLean
et al., 2021). Finally, the package integrates with the greater
tidymodels collection of packages, a burgeoning software ecosystem for
tidyverse-aligned predictive modeling used across many modern research
and industrial applications (Kuhn \& Wickham, 2020). To date, the
package has been downloaded more than 400,000 times.

@@

\begin{itemize}
\tightlist
\item
  \texttt{specify()} allows the analyst to specify the variable, or
  relationship between variables, that they are interested in.
\item
  \texttt{hypothesize()} allows the analyst to declare the null
  hypothesis.
\item
  \texttt{generate()} allows the analyst to generate data reflecting the
  null hypothesis or using the bootstrap.
\item
  \texttt{calculate()} allows the analyst to calculate summary
  statistics, either from

  \begin{itemize}
  \tightlist
  \item
    the observed data, to form the observed test statistic.
  \item
    data \texttt{generate()}d to reflect the null hypothesis, to form a
    randomization-based null distribution of test statistics.
  \end{itemize}
\end{itemize}

As such, the ultimate output of an infer pipeline using these four
functions is generally an \emph{observed statistic} or \emph{null
distribution} of test statistics. These four functions are thus
supplemented with several utilities to visualize and extract value from
their outputs.

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{shade\_p\_value()} situates the observed statistic in the
    null distribution, shading the region as or more extreme.
  \end{itemize}
\item
  \texttt{get\_p\_value()} calculates a p-value via the juxtaposition of
  the test statistic and null distribution.
\end{itemize}

The workflow outlined above can also be used for constructing confidence
intervals via bootstrapping with the omission of the
\texttt{hypothesize()} step in the pipeline. The resulting bootstrap
distribution can then be visualized with \texttt{visualize()}, the
confidence interval region can be situated in the bootstrap distribution
with \texttt{shade\_confidence\_interval()}, and the bounds of the
confidence interval can be calculated with
\texttt{get\_confidence\_interval()}.

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though do not provide methods for fitting models,
describing null distributions, performing bootstrapping, or calculating
summary statistics from tabular data. \texttt{statsexpressions}, and
adjacent packages in the \texttt{easystats} ecosystem, implement
wrappers with consistent interfaces for theory-based hypothesis tests.
At a higher level, the structure of each of these packages is defined by
model types, where each model type has its own associated function
and/or model object. In contrast, \texttt{infer} is structured around
four functions, situating statistics and model types within a more
abstracted grammar. \texttt{infer}'s initial release on the
Comprehensive R Archive Network predated that of each of the packages
mentioned above, with the exception of \texttt{broom}, a package
situated in the same \texttt{tidymodels} ecosystem (R Core Team, 2021).
